### 2022-09-12 manuscript/supplement.tex
add validation plot to supplement

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\begin{document}

\begin{center}
\textbf{Supplementary material for:}\\
\emph{Geographically aware estimates of remotely sensed water properties for Chesapeake Bay}
\end{center}

\begin{figure}[h!]
    \begin{center}
          \includegraphics[width=0.85\textwidth,keepaspectratio]{figures/counts_obs_table}
    \end{center}    
\end{figure}

\begin{figure}[h!]
    \begin{center}
          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_validation}
          \caption{RF predicted salinity versus test set (observed) salinity. Hatching indicates the area containing 80\% of the data points.}
    \end{center}    
\end{figure}

\clearpage

\end{document}