### 2022-09-19 manuscript/supplement.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\begin{document}

\begin{center}
\textbf{Supplementary material for:}\\
\emph{Geographically aware estimates of remotely sensed water properties for Chesapeake Bay}
\end{center}

\begin{figure}[h!]
    \begin{center}
          \includegraphics[width=0.85\textwidth,keepaspectratio]{figures/counts_obs_table}
    \end{center}    
\end{figure}

\begin{figure}[h!]
    \begin{center}
          \includegraphics[width=0.85\textwidth,keepaspectratio]{figures/obs_distribution_table}
    \end{center}    
\end{figure}

\begin{figure}[h!]
    \begin{center}
          \includegraphics[width=1\textwidth,keepaspectratio]{figures/rf_stats_table_0_}
    \end{center}    
\end{figure}

\begin{figure}[h!]
    \begin{center}
          \includegraphics[width=1\textwidth,keepaspectratio]{figures/_seasonality_turbidity}
          \caption{Seasonal Random Forest prediction results for turbidity.}
    \end{center}    
\end{figure}

\begin{figure}[h!]
    \begin{center}
          \includegraphics[width=1\textwidth,keepaspectratio]{figures/_freqcount_hex}
          \caption{Locations of observational data.}
    \end{center}    
\end{figure}

\clearpage

\end{document}